---
output:
  github_document:
    html_preview: false
output_file: readme.md
bibliography: sty/references.bib
csl: ./sty/apa-no-ampersand.csl
---

# X-vine models for multivariate extremes


The `Xvine` package is designed to analyze extremal dependence in multivariate extremes using the graphical structure of regular vines, as described in @kiriliouk2023x.
This modelling approach provides flexibility and enables exploration of sparsity.

Computationally, the package employs recursive approaches based on bivariate building blocks. The key components contain simulation from X-vine models, sequential parameter estimation, selection of bivariate parametric families, and vine structure selection.

##  Installation

The [GitHub](https://github.com/JeongjinLee88/Xvine.git) version can be downloaded via:

```{r, eval=FALSE}
#install.packages("devtools")
devtools::install_github("JeongjinLee88/Xvine")
```

##  Workflow

1. X-vine specification

```{r, eval=FALSE}
XVineSpec(M = StrMtx, Mmod = FamMtx, Mpar = ParMtx)
```

2. Simulation from Pareto distribution

```{r, eval=FALSE}
Dat_P=ParetoSim(n = 2000, XVS = XVS) # Pareto scale
```

3. Sequential parameter estimation via recursive approaches

```{r, eval=FALSE}
XVineSeqEst(data = Dat_P, Rank = T, qt = 0.05, XVS=XVS, method = 'mle')
```

4. Parametric family selection

```{r, eval=FALSE}

```

5. Vine structure selection

```{r, eval=FALSE}

```

## References