---
title: "XvineFlightDelays"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{XvineFlightDelays}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../inst/references.bib
---




```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

####    1. Load packages

```{r setup}
library(Xvine)
library(graphicalExtremes)
library(VineCopula)
library(ggplot2)
```


####    2. Load flight delays data

The raw data-set is accessible through the US Bureau of Transportation Statistics^[http://www.bts.dot.gov]. 
These flight delay data were analysed by @hentschel2022statistical who first selected airports with a minimum of $1000$ flights per year and applied a k-medoids clustering algorithm to identify homogeneous clusters in terms of extremal dependence.
Focusing on the H&udblac;sler–Reiss family, @hentschel2022statistical fitted an extremal graphical model to large flight delays of airports in the Texas cluster to investigate conditional independence.
Following the vignette in @engelke2graphicalextremes, we split the data set in half: a training set for model fitting and a test set for tuning parameter selection.
The dataset `FlightDelay` consists of the sample size of $n=3603$ and $d=29$ variables.

```{r, include=FALSE}
FlightTrain <- drop(graphicalExtremes::getFlightDelayData('delays', 'tcCluster', 'tcTrain'))
FlightTest <- drop(graphicalExtremes::getFlightDelayData('delays', 'tcCluster', 'tcTest'))
FlightDelay=rbind(FlightTrain,FlightTest)
#which(dimnames(FlightDelay)[[1]]>=2020)
#FlightDelay=FlightDelay[1:3522,] Exclude 2020
#summary(FlightDelay[3261:3522,])
#summary(FlightDelay[3523:3603,])
```


####    3. Fit the X-vine model to the whole data set.

Figure F.6. (a) shows a chi-plot comparing empirical pairwise tail dependence coefficients with those derived from the full X-vine model.
Figure F.6. (b) shows a plot comparing empirical trivariate tail dependence coefficients with those obtained from the fitted full X-vine model with $q^*=7$.


```{r, eval=FALSE}
# Time difference of 2.34063 mins
start <- Sys.time()
XVineFlightOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
              , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
              , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = FALSE
              , quan=0.13, tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
              , selectioncrit = "AIC", trunclevel = FALSE, progress = TRUE
              , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
              , tau_threshold = 0.05, weights = NA, cores=1)
print( Sys.time() - start )
XVineFlightOut$ChiLPlot # chi-plot
XVineFlightOut$mBIC_g # mBIC-plot
TriChiPlot=XVineFlightOut$Chi3Plot +
scale_x_continuous(n.breaks = 5,limits = c(0,0.8))+scale_y_continuous(limits=c(0,0.8))+theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm")) # trichi-plot
ggsave(TriChiPlot
, file="/Users/jlee/Desktop/XVine/TriChiFullXVine.pdf", width=4, height=4)

XVineFlightOut$VineTree[[1]]$E$Copula.type
sum(XVineFlightOut$VineTree[[1]]$E$Copula.type==3)
summary(XVineFlightOut$VineTree[[1]]$E$weights)
table(do.call(c,lapply(2:(29-1),function(i)(XVineFlightOut$VineTree[[i]]$E$Copula.type))))

save(XVineFlightOut,file = "/Users/jlee/Desktop/XVine/Rdata/XVineFlightOut_E2020.Rdata")
load(file = "/Users/jlee/Desktop/XVine/Rdata/XVineFlightOut_E2020.Rdata")

XVineFlightChiE2020=XVineFlightOut$ChiLPlot+
scale_x_continuous(n.breaks = 5,limits = c(0.15,0.8))+scale_y_continuous(limits=c(0.15,0.8))+theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))
XVineFlightmBICE2020=XVineFlightOut$mBIC_g
ggsave(XVineFlightmBICE2020, file="/Users/jlee/Desktop/XVine/XVineFlightmBICE2020.pdf", width=4, height=4)

```


####    4. Sensitive analysis for the mBIC-optimal truncation level across low quantiles.

Figure F.7 (b) shows a plot of the mBIC-optimal truncation level across a sequence of quantiles $k/n \in (0.05, 0.15)$.

```{r, eval=FALSE}
# Time difference of 15.89086 mins
start <- Sys.time()
mBICout=mBICplot(FlightDelay, Rank=TRUE, quan=seq(0.05,0.15,length.out=11)
         , famset_tc=c(1:4), famset_pc=c(0,1,3,4,5,6,13,14,16)
         , selectioncrit="AIC",trunclevel=FALSE
         , treecritT1="chi",treecritT2="tau"
         , effsampsize=10, tau_threshold=0.05, si=0.9)
print( Sys.time() - start )
ggsave(mBICout, file="/Users/jlee/Desktop/XVine/mBICsensitive.pdf", width=4, height=4)

```


####    5. Fit the X-vine model truncated at the mBIC-optimal truncation level

Figure 4. (e) shows a chi-plot comparing empirical pairwise tail dependence coefficients with those from the fitted truncated X-vine model.
Figure F.6 (c) shows a plot comparing empirical trivariate tail dependence coefficients with those obtained from the fitted truncated X-vine model with $q^*=7$.

```{r, eval=FALSE}
# Time difference of 58.30549 secs
XVineFlightTruncOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
              , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
              , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = FALSE
              , quan=0.13, tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
              , selectioncrit = "AIC", trunclevel = 7, progress = TRUE
              , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
              , tau_threshold = 0.05, weights = NA, cores=1)
XVineFlightTruncOut$ChiLPlot
TriChi_Trunc=XVineFlightTruncOut$Chi3Plot +
scale_x_continuous(n.breaks = 5,limits = c(0,0.8))+scale_y_continuous(limits=c(0,0.8))+theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm")) # trichi-plot
ggsave(TriChi_Trunc
, file="/Users/jlee/Desktop/XVine/TriChiTruncXVine.pdf", width=4, height=4)


# Except 2020
XVineFlightTruncOut=XVineModelFit(data = FlightDelay, N = 3500, XVS = NULL
              , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
              , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = FALSE
              , quan=0.13, tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
              , selectioncrit = "AIC", trunclevel = 8, progress = TRUE
              , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
              , tau_threshold = 0.05, weights = NA, cores=1)

XVineFlightTruncOut$VineTree[[1]]$E$Copula.type
sum(XVineFlightTruncOut$VineTree[[1]]$E$Copula.type==3)
summary(XVineFlightTruncOut$VineTree[[1]]$E$weights)
table(do.call(c,lapply(2:8,function(i)(XVineFlightTruncOut$VineTree[[i]]$E$Copula.type))))
XVineFlightTruncChiE2020=XVineFlightTruncOut$ChiLPlot+
scale_x_continuous(n.breaks = 5,limits = c(0.15,0.8))+scale_y_continuous(limits=c(0.15,0.8))+theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))

save(XVineFlightTruncOut,file = "/Users/jlee/Desktop/XVine/Rdata/XVineFlightTruncOut_E2020.Rdata")
ggsave(XVineFlightTruncChiE2020, file="/Users/jlee/Desktop/XVine/XVineFlightTruncChiE2020.pdf", width=4, height=4)

```


####    6. Fit the Husler-Reiss X-vine model at the mBIC-optimal truncation level


Figure F.7 (a) shows mBIC-values plotted across tree levels with a dotted line indicating a selected mBIC-optimal truncation level of $q^* = 7$.
Figure F.8 (b) shows a chi-plot for truncated HR X-vine with $q^*=7$ where we use the `Gamma2chi` function to transform the empirical variogram to the empirical chi.

```{r, eval=FALSE}
# Time difference of 45.66167 secs
# For the selection of T_1, used chi values as edge weight
start <- Sys.time()
HRXVineFlightOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
                , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
                , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = TRUE
                , MST1_HR=TRUE, quan=0.13, tcfamset = c(1), pcfamset = c(0,1)
                , selectioncrit = "AIC", trunclevel = 7, progress = TRUE
                , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
                , tau_threshold = 0.05, weights = NA, cores=1)
print( Sys.time() - start )

HRXVineFlightOut$ChiLPlot
HRXVineFlightOut$VarioPlot

# For the selection of T_1, used a empirical variogram as edge weight
HRXVineFlightOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
                , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
                , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = TRUE
                , MST1_HR=TRUE, quan=0.13, tcfamset = c(1), pcfamset = c(0,1)
                , selectioncrit = "AIC", trunclevel = FALSE, progress = TRUE
                , treecritT1 = "variogram", treecritT2 = "tau", effsampsize = 10
                , tau_threshold = 0.05, weights = NA, cores=1)
HRXVineFlightOut$VineTree[[19]]$E$Copula.type
table(do.call(c,lapply(2:28,function(i)(HRXVineFlightOut$VineTree[[i]]$E$Copula.type))))

HRXVineFlightOut$mBIC_g
HRXVineFlightOut$ChiLPlot
HRXVineFlightOut$VarioPlot
save(HRXVineFlightOut,file = "/Users/jlee/Desktop/XVine/Rdata/HRXVineFlightOut_E2020.Rdata")

```


```{r, eval=FALSE}
# For the selection of T_1, used a empirical variogram as edge weight
# Time difference of 45.68805 secs
HRXVineVarioFlightOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
                , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
                , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = TRUE
                , MST1_HR=TRUE, quan=0.13, tcfamset = c(1), pcfamset = c(0,1)
                , selectioncrit = "AIC", trunclevel = 7, progress = TRUE
                , treecritT1 = "variogram", treecritT2 = "tau", effsampsize = 10
                , tau_threshold = 0.05, weights = NA, cores=1)
HRXVineVarioFlightOut$ChiLPlot
HRXVineVarioFlightOut$VarioPlot+
scale_x_continuous(n.breaks = 5,limits = c(0.15,0.8))+scale_y_continuous(limits=c(0.15,0.8))+theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))

##  Exclude 2020
HRXVineVarioFlightOut=XVineModelFit(data = FlightDelay, N = 3500, XVS = NULL
                , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
                , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = TRUE
                , MST1_HR=TRUE, quan=0.13, tcfamset = c(1), pcfamset = c(0,1)
                , selectioncrit = "AIC", trunclevel = 6, progress = TRUE
                , treecritT1 = "variogram", treecritT2 = "tau", effsampsize = 10
                , tau_threshold = 0.05, weights = NA, cores=1)

HRXVineVarioFlightOut$VineTree[[1]]$E$Copula.type
sum(HRXVineVarioFlightOut$VineTree[[1]]$E$Copula.type==3)
summary(HRXVineVarioFlightOut$VineTree[[1]]$E$weights)
table(do.call(c,lapply(2:6,function(i)(HRXVineVarioFlightOut$VineTree[[i]]$E$Copula.type))))

save(HRXVineVarioFlightOut,file = "/Users/jlee/Desktop/XVine/Rdata/HRXVineVarioFlightOut_E2020.Rdata")
load(file = "/Users/jlee/Desktop/XVine/Rdata/HRXVineVarioFlightOut_E2020.Rdata")
HRXVineChiE2020=HRXVineVarioFlightOut$ChiLPlot+
scale_x_continuous(n.breaks = 5,limits = c(0.15,0.8))+scale_y_continuous(limits=c(0.15,0.8))+theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))
HRXVineVarioE2020=HRXVineVarioFlightOut$VarioPlot+
  scale_x_continuous(n.breaks = 5,limits = c(0.3,0.7))+scale_y_continuous(limits=c(0.3,0.7))+theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))+labs(title="",x=expression(paste("Empirical"," ",chi)), y = expression(paste("Fitted"," ",chi)))
ggsave(HRXVineChiE2020, file="/Users/jlee/Desktop/XVine/HRXVineChiE2020.pdf", width=4, height=4)
ggsave(HRXVineVarioE2020, file="/Users/jlee/Desktop/XVine/HRXVineVarioE2020.pdf", width=4, height=4)

```



####    7. Superimposed figure for the truncated X-vine model

This section is to reproduce Figure 4 (b): The X-vine graph with the first seven trees superimposed.

```{r, eval=FALSE}
FlightGraphSuperImposed=XVineFlightGraph(XVineOut = XVineFlightOut, TreeDimension = 7, IATAS = dimnames(FlightDelay)[[2]], Color4Edges = NULL)$ggp
```


####    8. Figures for an individual tree level by level for the X-vine model

This section is to reproduce Figure F.5: sequential display of subsequent trees of the truncated X-vine model.

```{r, eval=FALSE}
FlightGraphPerTree=XVineFlightGraphPerTree(XVineOut = XVineFlightOut, IATAS=dimnames(FlightDelay)[[2]], TreeLevel = 1, Color4Edge = NULL)$ggp
FlightGraphPerTree
```

Similarly, we can reproduce Figure F.9 by replacing the argument "XVineFlightOut" with "HRXVineVarioFlightOut".

```{r, eval=FALSE}
FlightGraphPerTree=XVineFlightGraphPerTree(XVineOut = HRXVineVarioFlightOut, IATAS=dimnames(FlightDelay)[[2]], TreeLevel = 2, Color4Edge = NULL)$ggp
FlightGraphPerTree
```

