---
title: "XvineFlightDelays"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{XvineFlightDelays}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---




```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

####    1. Load packages

```{r setup}
library(Xvine)
library(graphicalExtremes)
library(VineCopula)
library(ggplot2)
```


####    2. Load flight delays data

The raw data-set is accessible through the US Bureau of Transportation Statistics^[http://www.bts.dot.gov]. 
These flight delay data were analysed by @hentschel2022statistical who first selected airports with a minimum of 1 000 flights per year and applied a k-medoids clustering algorithm to identify homogeneous clusters in terms of extremal dependence.
Focusing on the H¨usler–Reiss family, @hentschel2022statistical fitted an extremal graphical model to large flight delays of airports in the Texas cluster to investigate conditional independence.
Following the vignette in Engelke et al. [2022a], we split the data set in half: a training set for model fitting and a test set for tuning parameter selection.
The dataset `FlightDelay` consists of the sample size of $n=3603$ and $d=29$ variables.

```{r, include=FALSE}
FlightTrain <- drop(graphicalExtremes::getFlightDelayData('delays', 'tcCluster', 'tcTrain'))
FlightTest <- drop(graphicalExtremes::getFlightDelayData('delays', 'tcCluster', 'tcTest'))
FlightDelay=rbind(FlightTrain,FlightTest)
```


####    3. Fit the X-vine model to the whole data set.

Fig 10. (a) shows a chi-plot comparing empirical pairwise tail dependence coefficients with those derived from the full X-vine model.
Fig 10. (b) shows a plot comparing empirical trivariate tail dependence coefficients with those obtained from the fitted full X-vine model with $q^*=7$.


```{r, eval=FALSE}
XVineFlightOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
              , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
              , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = FALSE
              , quan=0.13, tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
              , selectioncrit = "AIC", trunclevel = FALSE, progress = TRUE
              , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
              , tau_threshold = 0.05, weights = NA, cores=1)
XVineFlightOut$ChiLPlot
XVineFlightOut$mBIC_g
```

####    4. Sensitive analysis for the mBIC-optimal truncation level across low quantiles.

Fig 11. (b) shows a plot of the mBIC-optimal truncation level across a sequence of quantiles $k/n \in (0.05, 0.15)$.

```{r, eval=FALSE}
# Time difference of 15.89086 mins
start <- Sys.time()
mBICplot(FlightDelay, Rank=TRUE, quan=seq(0.05,0.15,length.out=11)
         , famset_tc=c(1:4), famset_pc=c(0,1,3,4,5,6,13,14,16)
         , selectioncrit="AIC",trunclevel=FALSE
         , treecritT1="chi",treecritT2="tau"
         , effsampsize=10, tau_threshold=0.05, si=0.9)
print( Sys.time() - start )
```


####    5. Fit the X-vine model truncated at the mBIC-optimal truncation level

Fig 5. (e) shows a chi-plot comparing empirical pairwise tail dependence coefficients with those from the fitted truncated X-vine model.


```{r, include=FALSE}
# Time difference of 58.30549 secs
XVineFlightTruncOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
              , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
              , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = FALSE
              , quan=0.13, tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
              , selectioncrit = "AIC", trunclevel = 7, progress = TRUE
              , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
              , tau_threshold = 0.05, weights = NA, cores=1)
XVineFlightTruncOut$ChiLPlot
```


####    6. Fit the Husler-Reiss X-vine model at the mBIC-optimal truncation level


Fig 10. (c) shows a plot comparing empirical trivariate tail dependence coefficients with those obtained from the fitted truncated X-vine model with $q^*=7$.
Fig 11. (a) shows mBIC-values plotted across tree levels with a dotted line indicating a selected mBIC-optimal truncation level of $q^* = 7$.
Fig 12. (b) shows a chi-plot for truncated HR X-vine with $q^*=7$.

```{r, include=FALSE}
# Time difference of 45.66167 secs
# For the selection of T_1, used chi values as edge weight
start <- Sys.time()
HRXVineFlightOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
                , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
                , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = TRUE
                , MST1_HR=TRUE, quan=0.13, tcfamset = c(1), pcfamset = c(0,1)
                , selectioncrit = "AIC", trunclevel = 7, progress = TRUE
                , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
                , tau_threshold = 0.05, weights = NA, cores=1)
print( Sys.time() - start )

HRXVineFlightOut$ChiLPlot
HRXVineFlightOut$VarioPlot
```


```{r, include=FALSE}
# For the selection of T_1, used a empirical variogram as edge weight
# Time difference of 45.68805 secs
HRXVineVarioFlightOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
                , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
                , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = TRUE
                , MST1_HR=TRUE, quan=0.13, tcfamset = c(1), pcfamset = c(0,1)
                , selectioncrit = "AIC", trunclevel = 7, progress = TRUE
                , treecritT1 = "variogram", treecritT2 = "tau", effsampsize = 10
                , tau_threshold = 0.05, weights = NA, cores=1)
HRXVineVarioFlightOut$ChiLPlot
HRXVineVarioFlightOut$VarioPlot+
scale_x_continuous(n.breaks = 5,limits = c(0.15,0.8))+scale_y_continuous(limits=c(0.15,0.8))+theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))
```



####    7. Superimposed figure for the truncated X-vine model

This section is to reproduce Fig 5. (b): The X-vine graph with the first seven trees superimposed.

```{r, eval=FALSE}
FlightGraphSuperImposed=XVineFlightGraph(XVineOut = XVineFlightOut, TreeDimension = 7, IATAS = dimnames(FlightDelay)[[2]], Color4Edges = NULL)$ggp
```


####    8. Figures for an individual tree level by level for the X-vine model

This section is to reproduce Fig. 9: sequential display of subsequent trees of the truncated X-vine model.

```{r, eval=FALSE}
FlightGraphPerTree=XVineFlightGraphPerTree(XVineOut = XVineFlightOut, IATAS=dimnames(FlightDelay)[[2]], TreeLevel = 1, Color4Edge = NULL)$ggp
FlightGraphPerTree
```

Similarly, we can reproduce Fig. 13 by replacing the argument "XVineFlightOut" with "HRXVineVarioFlightOut".

```{r, eval=FALSE}
FlightGraphPerTree=XVineFlightGraphPerTree(XVineOut = HRXVineVarioFlightOut, IATAS=dimnames(FlightDelay)[[2]], TreeLevel = 2, Color4Edge = NULL)$ggp
FlightGraphPerTree
```
