---
title: "XvineFlightDelays"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{XvineFlightDelays}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---




```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

####    1. Load packages

```{r setup}
library(Xvine)
library(graphicalExtremes)
library(VineCopula)
```


####    2. Load flight delays data

The flight delays dataset consists of the sample size of $n=3603$ and $d=29$ variables.

```{r, include=FALSE}
FlightTrain <- drop(graphicalExtremes::getFlightDelayData('delays', 'tcCluster', 'tcTrain'))
FlightTest <- drop(graphicalExtremes::getFlightDelayData('delays', 'tcCluster', 'tcTest'))
FlightDelay=rbind(FlightTrain,FlightTest)
```


####    3. Fit the X-vine model to the whole data set.


```{r, eval=FALSE}
XVineFlightOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
              , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
              , Chi3_graph = TRUE, Rank_Vario = FALSE, Vario_graph = FALSE
              , quan=0.13, tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
              , selectioncrit = "AIC", trunclevel = FALSE, progress = TRUE
              , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
              , tau_threshold = 0.05, weights = NA, cores=1)
XVineFlightOut$ChiLPlot
XVineFlightOut$mBIC_g
```


####    4. Fit the X-vine model truncated at the mBIC-optimal truncation level


```{r, include=FALSE}
XVineFlightTruncOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
              , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
              , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = FALSE
              , quan=0.13, tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
              , selectioncrit = "AIC", trunclevel = 7, progress = TRUE
              , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
              , tau_threshold = 0.05, weights = NA, cores=1)
XVineFlightTruncOut$ChiLPlot
```


####    5. Fit the Husler-Reiss X-vine model at the mBIC-optimal truncation level


```{r, include=FALSE}
HRXVineFlightOut=XVineModelFit(data = FlightDelay, N = 3600, XVS = NULL
                , Rank = TRUE, Rank_chiL = TRUE, Rank_chi3 = TRUE
                , Chi3_graph = TRUE, Rank_Vario = TRUE, Vario_graph = TRUE
                , MST1_HR=TRUE, quan=0.13, tcfamset = c(1), pcfamset = c(0,1)
                , selectioncrit = "AIC", trunclevel = 7, progress = TRUE
                , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
                , tau_threshold = 0.05, weights = NA, cores=1)
HRXVineFlightOut$ChiLPlot
HRXVineFlightOut$VarioPlot
```


####    6. Sensitive analysis for the mBIC-optimal truncation level across low quantiles.


```{r, eval=FALSE}
mBICplot(FlightDelay, Rank=TRUE, quan=seq(0.05,0.15,by=0.1)
         , famset_tc=c(1:4), famset_pc=c(0,1,3,4,5,6,13,14,16)
         , selectioncrit="AIC",trunclevel=FALSE
         , treecritT1="chi",treecritT2="tau"
         , effsampsize=10, tau_threshold=0.05, si=0.9)
```


####    7. Superimposed figure for the truncated X-vine model

```{r, eval=FALSE}
FlightGraphSuperImposed=XVineFlightGraph(XVineOut = XVineOut_Flight, TreeDimension = 7, IATAS = dimnames(FlightDelay)[[2]], Color4Edges = NULL)$ggp
```


####    8. Figures for an individual tree level by level for the X-vine model

```{r, eval=FALSE}
FlightGraphPerTree=XVineFlightGraphPerTree(XVineOut = XVineOut_Flight, IATAS=dimnames(FlightDelay)[[2]], TreeLevel = 2, Color4Edge = NULL)$ggp
```

