---
title: "XvineReproducibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{XvineReproducibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---




```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

###   1. Load required packages.

The X-Vine package requires other packages: `VineCopula` and `graphicalExtremes`.

```{r, echo=TRUE,message=FALSE,warning=FALSE}
library(Xvine)
library(VineCopula)
library(ggplot2)
```


###   2. X-Vine specification.

`XVineSpec()` function specifies the X-Vine specification $(\mathcal{V},\mathcal{R},\mathcal{C})$; vine tree structure matrix, family matrix for bivariate parametric (tail) copula families, and the associated parameter matrix. 

```{r, include=FALSE}
StrMtx <- matrix(c(1, 1, 2, 2, 4,
                   0, 2, 1, 3, 2,
                   0, 0, 3, 1, 3,
                   0, 0, 0, 4, 1,
                   0, 0, 0, 0, 5),5,byrow = T)
ParMtx <- matrix(c(0, 1.5, 2, 2.5, 2,
                   0, 0, 2, 2.5, 0.7,
                   0, 0, 0, 0.4, -0.3,
                   0, 0, 0, 0, 0.1,
                   0, 0, 0, 0, 0),5,byrow = T)
FamMtx <- matrix(c(0, 1, 2, 3, 4,
                   0, 0, 3, 4, 1,
                   0, 0, 0, 3, 1,
                   0, 0, 0, 0, 1,
                   0, 0, 0, 0, 0),5,byrow = T)
XVS=XVineSpec(M = StrMtx, Mmod = FamMtx, Mpar = ParMtx)
```


###   3. Create box-plots of parameter estimates for X-vine models.

This section is to reproduce Fig 3. (a) and (b), and Fig7.

Given the X-Vine structure and selected bivariate parametric families, we draw multivariate (generalized) Pareto samples of size $N=4000$ and set a threshold $k/n=0.05$ of the uniform margin $\widehat{U}_{i,j}$
Instead of using these Pareto samples directly, we perform a rank transformation and take sub-samples to investigate results with respect to threshold exceedances.
We obtain ML estimates under 200 repeated iterations.
We also convert the ML estimates to dependence measure coefficients: the tail dependence coefficient for $e\in E_1$ and Kendall's tau for $e\in E_i$, $i>1$ using the parametric relationships.
Fig 3. (a) shows box-plots of dependence measure estimates from specified bivairate parametric families.
Varying $(n,k)$, we can reproduce box-plots in Fig. 7.
For Fig 3. (b), bivariate parametric (tail) families are selected sequentially from the data along the vine tree sequence.


```{r, include=FALSE}
# n = 1000, 2000, 4000 / k/n = 0.02, 0.05, 0.1
# Time difference of 1.068745 mins
BoxOut4K05=XVineBoxplot(N = 4000, qt = 0.05, ite = 200, XVS = XVS, RankT = TRUE)

d=5 # the number of variables
TrueMtx_dep=t(DependMtx(XVS)) # Dependence measures
y0s <- c(TrueMtx_dep[lower.tri(TrueMtx_dep)])
dLines <- data.frame(X =1:choose(d,2) - 0.4,
                     Y = y0s,
                     Xend = 1:choose(d,2) + 0.4,
                     Yend = y0s,
                     Group = c("typeA"),
                     color = c("red"))
  
BP_DEP_SPE <- BoxOut4K05$bp_dep_SPE +
scale_fill_manual(labels=c(expression(T[1]),expression(T[2]),expression(T[3]),expression(T[4])),values=c("lightskyblue","lightgreen","lavender","papayawhip")) +
theme(legend.title = element_blank()
     ,legend.text = element_text(face="bold",size=15)
     ,legend.key.size = unit(1, 'cm')
     ,legend.position="none"
     ,aspect.ratio = 1
     ,axis.text.x = element_text(color="black", face="bold", size=9, angle = 30, hjust = 0.5)
     ,axis.text.y = element_text(face="bold", size=9)
     ,panel.background = element_rect(fill = "white"
     ,colour = "white"
     ,linewidth = 0.5, linetype = "solid")
     ,panel.grid.major = element_line(linewidth = 0.5, linetype = 'solid',
                                                        colour = "grey")
     ,panel.grid.minor = element_line(linewidth = 0.25, linetype = 'solid',
                                                        colour = "grey")
     ,plot.margin = margin(0, 1, 0, 0, "cm")) +
scale_x_discrete(labels=c(expression(atop(chi[12],"\n(HR)")),expression(atop(chi[23],"\n(NL)")),expression(atop(chi[24],"\n(L)")),expression(atop(chi[45],"\n(Diri)")),expression(atop(tau[13~";"~2],"\n(Clay)")),expression(atop(tau[34~";"~2],"\n(Gum)")),expression(atop(tau[25~";"~4],"\n(Ga)")),expression(atop(tau[14~";"~23],"\n(Clay)")),expression(atop(tau[35~";"~24],"\n(Ga)")),expression(atop(tau[15~";"~234],"\n(Ga)")))) +
scale_y_continuous(breaks = seq(-0.4,0.8,0.2),limits = c(-0.4,0.8)) +
geom_segment(data = dLines, color = "red",aes(x = X, y = Y, xend = Xend, yend=Yend), inherit.aes = FALSE)
BP_DEP_SPE

BP_DEP_FAM <- BoxOut4K05$bp_dep_Fam +
scale_fill_manual(labels=c(expression(T[1]),expression(T[2]),expression(T[3]),expression(T[4])),values=c("lightskyblue","lightgreen","lavender","papayawhip")) +
theme(legend.title = element_blank()
     ,legend.text = element_text(face="bold",size=15)
     ,legend.key.size = unit(1, 'cm')
     ,legend.position="none"
     ,aspect.ratio = 1
     ,axis.text.x = element_text(color="black", face="bold", size=9, angle = 30, hjust = 0.5)
     ,axis.text.y = element_text(face="bold", size=9)
     ,panel.background = element_rect(fill = "white"
     ,colour = "white"
     ,linewidth = 0.5, linetype = "solid")
     ,panel.grid.major = element_line(linewidth = 0.5, linetype = 'solid',
                                                        colour = "grey")
     ,panel.grid.minor = element_line(linewidth = 0.25, linetype = 'solid',
                                                        colour = "grey")
     ,plot.margin = margin(0, 1, 0, 0, "cm")) +
geom_segment(data = dLines, color = "red",aes(x = X, y = Y, xend = Xend, yend=Yend), inherit.aes = FALSE) + scale_x_discrete(labels=c(expression(chi[12]),expression(chi[23]),expression(chi[24]),expression(chi[45]),expression(tau[13~";"~2]),expression(tau[34~";"~2]),expression(tau[25~";"~4]),expression(tau[14~";"~23]),expression(tau[35~";"~24]),expression(tau[15~";"~234]))) +
scale_y_continuous(breaks = seq(-0.4,0.8,0.2),limits = c(-0.4,0.8))
BP_DEP_FAM
```


####    4. Box-plots of ML estimates for the tail copula density linked to a specific edge.

This section is to reproduce Fig 8.
`XVineBoxplot()` creates box-plots of parameter estimates for a parameter ($\Gamma$ or $\theta$, etc.) corresponding a specific edge.
In Fig 8., we consider the box-plots of ML estimates for the H\"{u}sler-Reiss tail copula density linked to the edge $e=\{1,2\}\in E_1$, using different $(n,k)$.

```{r, eval=FALSE}
# n = 1000, 2000, 4000 / k/n = 0.02, 0.05, 0.1
ite=200
BoxOut1K02=XVineBoxplot(N = 1000, qt = 0.02, ite = ite, XVS = XVS, RankT = T)
BoxOut1K05=XVineBoxplot(N = 1000, qt = 0.05, ite = ite, XVS = XVS, RankT = T)
BoxOut1K1=XVineBoxplot(N = 1000, qt = 0.1, ite = ite, XVS = XVS, RankT = T)

BoxPar1K=XVineBoxPlotEachPar(BoxOut1 = BoxOut1K02$MLE_WithoutFamSel,BoxOut2 = BoxOut1K05$MLE_WithoutFamSel,BoxOut3 = BoxOut1K1$MLE_WithoutFamSel,
                             RowInd = 1,ColInd = 2,XVS = XVS)
BoxPar1K$BoxOnePar+labs(title="",x="", y = "") +
scale_x_discrete(labels=c("k=20","k=50","k=100"))
```


####    5. Simulation for the selection of bivariate parametric families

This section is to reproduce Table 1. (a) and (b).
`XVineFamSelSimul()` returns simulation results of selecting bivariate parametric families by calling the function `XVineFamSel()`.
All information about the simulation result is saved in the list of `FamSelOut`.


```{r, eval=FALSE}
# Time difference of 1.115222 mins
FamSelOut <- XVineFamSelSimul(N = 4000, qt = 0.05, ite = 200, XVS = XVS
 ,edgeIndex = c(4,5), familyset_tc=c(1:4), familyset_cop = c(0,1,3,4,5,6,13,14,16))
FamSelOut$EdgeProp
FamSelOut$EffSizePercent
FamSelOut$GlobalProp
```


### 5. 50 dimensional X-vine specification 

Now, we consider a 50-dimensional C-vine where the first tree includes the
H\"{u}sler–Reiss models and negative logistic models with randomly assigned parameter values $\theta_e \in [1, 2]$ for $e \in E_1$. 
Subsequent trees contain bivariate Gaussian copulas with partial correlations $\rho_e = 1.1 − 0.1j$ for $e \in E_j$ and $j \in \{2, . . . , 9\}$, and $\rho_e = 0.1$ for $e \in E_j$ with $j \ge 10$.

```{r, eval=FALSE}
Dim=50
StrMtx=matrix(0,nrow = Dim,ncol = Dim)
for(i in 1:Dim){
  for(j in i:Dim){
    StrMtx[i,j]=i # C-vine
  }
}
set.seed(1220)
FamMtx=matrix(0,Dim,Dim)
for(i in 1:Dim){
  for(j in i:Dim){
    if(i==1){
    FamMtx[1,j]=sample(c(1,2),size = 1,replace = T) # assign randomly HR and NL in T1
    }else{
    FamMtx[i,j]=1 # assign all Gaussian copulas in Ti, i>1
    }
  }
}
diag(FamMtx)=0
set.seed(1220)
ParMtx=matrix(0,nrow = Dim,ncol = Dim)
for(i in 1:Dim){ 
  for(j in i:Dim){
    if(i==1){
      ParMtx[1,j]=runif(1,1,2)  # Parameter values on (1,2)
    }
    if(1 < i & i <= 9){
      ParMtx[i,j]=1.1-0.1*i
    }
    if(i > 9){
      ParMtx[i,j]=0.1 # same partial correlation 0.5 for Ti
    }
  }
}
diag(ParMtx)=0
##  X-Vine specification (Vine structure, Bivariate parametric families, Parameters)
# Time difference of 1.115222 mins
XVS=XVineSpec(M = StrMtx, Mmod = FamMtx, Mpar = ParMtx)
```


###   6. Fitting X-vine models.

This section is to reproduce Fig 4.
Now, we estimate all three components of the X-vine specification $(\mathcal{V},\mathcal{R},\mathcal{C})$.
We draw multivariate Pareto samples of size $n=1\,000$. 
Using the (lower) tail dependence coefficient $\hat{\chi}_e$ for $e\in E_1$ as edge weight and the Kendall's tau $\hat{\tau}_e$ for $e\in E_i$, $i\ge 2$, we first estimate the vine tree sequence sequentially from $T_1$ to $T_{d-1}$ that satisfies the proximity condition.
For the purpose of capturing as much extremal dependence as possible in the first several trees, we find maximum spanning trees.
After the vine tree sequence is selected, we choose bivariate parametric families for each edge in each tree and estimate the associated parameter(s).
Fig 4. (a) The mBIC plotted across tree levels with a dotted line indicating the selected mBIC-optimal truncation level of $q^∗ = 19$.
Fig 4. (b) $\chi$-plot comparing pairwise $\chi$ values from the specified X-vine model to those obtained from the fitted X-vine model via Monte Carlo simulations. Fig 4. (c) $\chi$-plot for the truncated X-vine model.

```{r, eval=FALSE}
# 'XVinePa50dim.Rdata' is included in the Xvine package.
# Time difference of 7.086743 mins
#set.seed(1953)
#XVinePa50dim=ParetoSim(n = 1000, XVS = XVS) # Pareto scale

#Time difference of 26.23817 mins
XVineFitOut=XVineModelFit(data = 1/XVinePa50dim, N = 5000, XVS = XVS
          , Rank = FALSE, Rank_chiL = FALSE, Rank_chi3 = FALSE
          , Chi3_graph = FALSE, Rank_Vario = FALSE, Vario_graph = FALSE
          , tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
          , selectioncrit = "AIC", trunclevel = FALSE, progress = TRUE
          , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
          , tau_threshold = 0.05, weights = NA, cores=1)
XVineFitOut$mBIC_g
XVineFitOut$ChiLPlot


XVineTrunOut=XVineModelFit(data = 1/XVinePa50dim, N = 5000, XVS = XVS
          , Rank = FALSE, Rank_chiL = FALSE, Rank_chi3 = FALSE
          , Chi3_graph = FALSE, Rank_Vario = FALSE, Vario_graph = FALSE
          , tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
          , selectioncrit = "AIC", trunclevel = 19, progress = TRUE
          , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
          , tau_threshold = 0.05, weights = NA, cores=1)
# Time difference of 18.49413 mins
XVineTrunOut$ChiLPlot
```

```{r, eval=FALSE}
XVineFitOut$ChiLPlot +
scale_x_continuous(breaks = seq(0.5,1,by=0.1),limits = c(0.45,1)) +
scale_y_continuous(breaks = seq(0.5,1,by=0.1),limits = c(0.45,1))

XVineTrunOut$ChiLPlot +
scale_x_continuous(breaks = seq(0.5,1,by=0.1),limits = c(0.45,1)) +
scale_y_continuous(breaks = seq(0.5,1,by=0.1),limits = c(0.45,1))

XVineFitOut$mBIC_g
```







