---
title: "XvineReproducibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{XvineReproducibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---




```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### 1. Load required packages.

The X-Vine package requires other packages: `VineCopula` and `graphicalExtremes`.

```{r, echo=TRUE,message=FALSE,warning=FALSE}
library(Xvine)
library(VineCopula)
```


### 2. X-Vine specification.

`XVineSpec()` function specifies the X-Vine specification $(\mathcal{V},\mathcal{R},\mathcal{C})$; vine tree structure matrix, family matrix for bivariate parametric (tail) copula families, and the associated parameter matrix. 

```{r, include=FALSE}
StrMtx <- matrix(c(1, 1, 2, 2, 4,
                   0, 2, 1, 3, 2,
                   0, 0, 3, 1, 3,
                   0, 0, 0, 4, 1,
                   0, 0, 0, 0, 5),5,byrow = T)
ParMtx <- matrix(c(0, 1.5, 2, 2.5, 2,
                   0, 0, 2, 2.5, 0.7,
                   0, 0, 0, 0.4, -0.3,
                   0, 0, 0, 0, 0.1,
                   0, 0, 0, 0, 0),5,byrow = T)
FamMtx <- matrix(c(0, 1, 2, 3, 4,
                   0, 0, 3, 4, 1,
                   0, 0, 0, 3, 1,
                   0, 0, 0, 0, 1,
                   0, 0, 0, 0, 0),5,byrow = T)
XVS=XVineSpec(M = StrMtx, Mmod = FamMtx, Mpar = ParMtx)
```


### 3. Create box-plots of ML estimates in terms of dependence measure coefficients, using different (n,k).

`XVineBoxPlot()` creates box-plots of parameter estimates of dependence measure coefficients for X-vine models. 
In Fig 3. (a), we are given the X-Vine structure and selected bivariate parametric families. 
We draw multivariate Pareto random samples of size $N=4000$.
Instead of using these Pareto samples directly, we perform a rank transformation and take sub-samples to investigate results with respect to threshold exceedances.
Setting a threshold $k/n=0.05$ of the uniform margin $\widehat{U}_{i,j}$, we obtain ML estimates under 200 repeated iterations.
We also convert the ML estimates to dependence measure coefficients: the tail dependence coefficient for $e\in E_1$ and Kendall's tau for $e\in E_i$, $i>1$ using the parametric relationships.
In Fig 3. (b), we select the families of bivariate parametric (tail) copulas along the vine tree sequence.

```{r, include=FALSE}
# n = 1000, 2000, 4000 / k/n = 0.02, 0.05, 0.1
BoxOut4K05=XVineBoxplot(N = 4000, qt = 0.05, ite = 200, XVS = XVS, RankT = TRUE)

d=5 # the number of variables
TrueMtx_dep=t(DependMtx(XVS)) # Dependence measures
y0s <- c(TrueMtx_dep[lower.tri(TrueMtx_dep)])
dLines <- data.frame(X =1:choose(d,2) - 0.4,
                     Y = y0s,
                     Xend = 1:choose(d,2) + 0.4,
                     Yend = y0s,
                     Group = c("typeA"),
                     color = c("red"))
  
BP_DEP_SPE <- BoxOut4K05$bp_dep_SPE +
scale_fill_manual(labels=c(expression(T[1]),expression(T[2]),expression(T[3]),expression(T[4])),values=c("lightskyblue","lightgreen","lavender","papayawhip")) +
theme(legend.title = element_blank()
     ,legend.text = element_text(face="bold",size=15)
     ,legend.key.size = unit(1, 'cm')
     ,legend.position="none"
     ,aspect.ratio = 1
     ,axis.text.x = element_text(color="black", face="bold", size=9, angle = 30, hjust = 0.5)
     ,axis.text.y = element_text(face="bold", size=9)
     ,panel.background = element_rect(fill = "white"
     ,colour = "white"
     ,linewidth = 0.5, linetype = "solid")
     ,panel.grid.major = element_line(linewidth = 0.5, linetype = 'solid',
                                                        colour = "grey")
     ,panel.grid.minor = element_line(linewidth = 0.25, linetype = 'solid',
                                                        colour = "grey")
     ,plot.margin = margin(0, 1, 0, 0, "cm")) +
scale_x_discrete(labels=c(expression(atop(chi[12],"\n(HR)")),expression(atop(chi[23],"\n(NL)")),expression(atop(chi[24],"\n(L)")),expression(atop(chi[45],"\n(Diri)")),expression(atop(tau[13~";"~2],"\n(Clay)")),expression(atop(tau[34~";"~2],"\n(Gum)")),expression(atop(tau[25~";"~4],"\n(Ga)")),expression(atop(tau[14~";"~23],"\n(Clay)")),expression(atop(tau[35~";"~24],"\n(Ga)")),expression(atop(tau[15~";"~234],"\n(Ga)")))) +
scale_y_continuous(breaks = seq(-0.4,0.8,0.2),limits = c(-0.4,0.8)) +
geom_segment(data = dLines, color = "red",aes(x = X, y = Y, xend = Xend, yend=Yend), inherit.aes = FALSE)
BP_DEP_SPE

BP_DEP_FAM <- BoxOut4K05$bp_dep_Fam +
scale_fill_manual(labels=c(expression(T[1]),expression(T[2]),expression(T[3]),expression(T[4])),values=c("lightskyblue","lightgreen","lavender","papayawhip")) +
theme(legend.title = element_blank()
     ,legend.text = element_text(face="bold",size=15)
     ,legend.key.size = unit(1, 'cm')
     ,legend.position="none"
     ,aspect.ratio = 1
     ,axis.text.x = element_text(color="black", face="bold", size=9, angle = 30, hjust = 0.5)
     ,axis.text.y = element_text(face="bold", size=9)
     ,panel.background = element_rect(fill = "white"
     ,colour = "white"
     ,linewidth = 0.5, linetype = "solid")
     ,panel.grid.major = element_line(linewidth = 0.5, linetype = 'solid',
                                                        colour = "grey")
     ,panel.grid.minor = element_line(linewidth = 0.25, linetype = 'solid',
                                                        colour = "grey")
     ,plot.margin = margin(0, 1, 0, 0, "cm")) +
geom_segment(data = dLines, color = "red",aes(x = X, y = Y, xend = Xend, yend=Yend), inherit.aes = FALSE) + scale_x_discrete(labels=c(expression(chi[12]),expression(chi[23]),expression(chi[24]),expression(chi[45]),expression(tau[13~";"~2]),expression(tau[34~";"~2]),expression(tau[25~";"~4]),expression(tau[14~";"~23]),expression(tau[35~";"~24]),expression(tau[15~";"~234]))) +
scale_y_continuous(breaks = seq(-0.4,0.8,0.2),limits = c(-0.4,0.8))
BP_DEP_FAM
```

#### 4. Simulation for the selection of bivariate parametric families

`XVineFamSelSimul` returns simulation results of selecting bivariate parametric families by calling the function [XVineFamSel()].
All information about the simulation result is saved in the list of `FamSelOut`.
Table 1. (a) and (b)

```{r, include=FALSE}
start <- Sys.time()
FamSelOut <- XVineFamSelSimul(N = 4000, qt = 0.05, ite = 200, XVS = XVS
 ,edgeIndex = c(4,5), familyset_tc=c(1:4), familyset_cop = c(0,1,3,4,5,6,13,14,16))
print( Sys.time() - start )
FamSelOut$EdgeProp
FamSelOut$EffSizePercent
```


### 5. 50 dimensional X-vine specification 


```{r, eval=FALSE}
##  2. A 50-dim X-vine model where HR and NL in T1 and Gaussian copulas in Ti, i>1
Dim=50
StrMtx=matrix(0,nrow = Dim,ncol = Dim)
for(i in 1:Dim){
  for(j in i:Dim){
    StrMtx[i,j]=i # C-vine
  }
}
set.seed(1220)
FamMtx=matrix(0,Dim,Dim)
for(i in 1:Dim){
  for(j in i:Dim){
    if(i==1){
    FamMtx[1,j]=sample(c(1,2),size = 1,replace = T) # assign randomly HR and NL in T1
    }else{
    FamMtx[i,j]=1 # assign all Gaussian copulas in Ti, i>1
    }
  }
}
diag(FamMtx)=0
set.seed(1220)
ParMtx=matrix(0,nrow = Dim,ncol = Dim)
for(i in 1:Dim){ 
  for(j in i:Dim){
    if(i==1){
      ParMtx[1,j]=runif(1,1,2)  # Parameter values on (1,2)
    }
    if(1 < i & i <= 9){
      ParMtx[i,j]=1.1-0.1*i
    }
    if(i > 9){
      ParMtx[i,j]=0.1 # same partial correlation 0.5 for Ti
    }
  }
}
diag(ParMtx)=0
##  X-Vine specification (Vine structure, Bivariate parametric families, Parameters)
XVS=XVineSpec(M = StrMtx, Mmod = FamMtx, Mpar = ParMtx)
```


### 6. Fitting X-vine models.

We estimate all three components of the X-vine specification $(\mathcal{V},\mathcal{R},\mathcal{C})$.
We draw multivariate Pareto samples of size $n=1\,000$. 
Using the (lower) tail dependence coefficient $\hat{\chi}_e$ for $e\in E_1$ as edge weight and the Kendall's tau $\hat{\tau}_e$ for $e\in E_i$, $i\ge 2$, we first estimate the vine tree sequence sequentially from $\T_1$ to $\T_{d-1}$ that satisfies the proximity condition.
For the purpose of capturing as much extremal dependence as possible in the first several trees, we find maximum spanning trees.
After the vine tree sequence is selected, we choose bivariate parametric families for each edge in each tree and estimate the associated parameter(s).

Fig. 4 (a) mBIC (b) chi-plot for a full model (c) chi-plot for truncated X-vine model


```{r,include=FALSE}
set.seed(1953)
Dat_P=ParetoSim(n = 1000, XVS = XVS) # Pareto scale
XvinePa50dim=Dat_P

start <- Sys.time()
XVineFitOut=XVineModelFit(data = 1/Dat_P, N = 5000, XVS = XVS
          , Rank = FALSE, Rank_chiL = FALSE, Rank_chi3 = FALSE
          , Chi3_graph = FALSE, Rank_Vario = FALSE, Vario_graph = FALSE
          , tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
          , selectioncrit = "AIC", trunclevel = FALSE, progress = TRUE
          , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
          , tau_threshold = 0.05, weights = NA, cores=1)
print( Sys.time() - start )
#Time difference of 26.23817 mins
XVineFitOut$mBIC_g
XVineFitOut$ChiLPlot

start <- Sys.time()
XVineTrunOut=XVineModelFit(data = 1/Dat_P, N = 5000, XVS = XVS
          , Rank = FALSE, Rank_chiL = FALSE, Rank_chi3 = FALSE
          , Chi3_graph = FALSE, Rank_Vario = FALSE, Vario_graph = FALSE
          , tcfamset = c(1:4), pcfamset = c(0,1,3,4,5,6,13,14,16)
          , selectioncrit = "AIC", trunclevel = 19, progress = TRUE
          , treecritT1 = "chi", treecritT2 = "tau", effsampsize = 10
          , tau_threshold = 0.05, weights = NA, cores=1)
print( Sys.time() - start )
# Time difference of 18.49413 mins
XVineTrunOut$ChiLPlot
```

```{r,eval=FALSE}
XVineFitOut$ChiLPlot +
scale_x_continuous(breaks = seq(0.5,1,by=0.1),limits = c(0.45,1)) +
scale_y_continuous(breaks = seq(0.5,1,by=0.1),limits = c(0.45,1))

XVineTrunOut$ChiLPlot +
scale_x_continuous(breaks = seq(0.5,1,by=0.1),limits = c(0.45,1)) +
scale_y_continuous(breaks = seq(0.5,1,by=0.1),limits = c(0.45,1))

XVineFitOut$mBIC_g
```




