---
title: "XvineWorkflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{XvineWorkflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---




```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### 1. Load required packages.

The X-Vine package requires other packages: `VineCopula` and `graphicalExtremes`.

```{r, echo=TRUE,message=FALSE,warning=FALSE}
library(Xvine)
library(VineCopula)
```


### 2. X-Vine specification.

The X-Vine specification $(\mathcal{V},\mathcal{R},\mathcal{C})$ consists of three components; vine tree structure matrix, family matrix for bivariate parametric (tail) copula families, and the associated parameter matrix. 

```{r, include=FALSE}
StrMtx <- matrix(c(1, 1, 2, 2, 4,
                   0, 2, 1, 3, 2,
                   0, 0, 3, 1, 3,
                   0, 0, 0, 4, 1,
                   0, 0, 0, 0, 5),5,byrow = T)
ParMtx <- matrix(c(0, 1.5, 2, 2.5, 2,
                   0, 0, 2, 2.5, 0.7,
                   0, 0, 0, 0.4, -0.3,
                   0, 0, 0, 0, 0.1,
                   0, 0, 0, 0, 0),5,byrow = T)
FamMtx <- matrix(c(0, 1, 2, 3, 4,
                   0, 0, 3, 4, 1,
                   0, 0, 0, 3, 1,
                   0, 0, 0, 0, 1,
                   0, 0, 0, 0, 0),5,byrow = T)
XVS=XVineSpec(M = StrMtx, Mmod = FamMtx, Mpar = ParMtx)
```


### 3. Simulation from X-Vine models.

Given the X-Vine specification $(\mathcal{V},\mathcal{R},\mathcal{C})$, the function `XVineSim` generate random samples from $\boldsymbol{Z}^{(j)}\sim R(\boldsymbol{z})\mathbb{I}(Z^{(j)}<1)$ for each $j=1,\ldots,d$.

```{r, include=FALSE}
XVineSim(N = 100, XVS = XVS, k = 1)
```

### 4. Simulation from multivariate Pareto distribution.

To simulate from the X-vine multivariate Pareto distribution $\boldsymbol{Y}_1,\ldots,\boldsymbol{Y}_n$, the function `ParetoSim` follows the exact simulation for multivariate Pareto distribution $\boldsymbol{Y}$, using the X-vine specification.
We can easily switch multivariate Pareto samples to inverted-Pareto samples by taking the reciprocal of $\boldsymbol{Z}_i=1/\boldsymbol{Y}_i$.

```{r, include=FALSE}
Dat_P=ParetoSim(n = 5000, XVS = XVS) # Pareto scale
```


### 5. Sequential Parameter Estimation for X-Vine models.

In order to implement the sequential parameter estimation, it is sufficient to use the structure matrix where diagonal elements are put in increasing order.
That is, we use the first structure matrix from permuted structure matrices and define its corresponding max matrix to find the appropriate arguments for conditional bivariate copula densities.

Given the vine tree structure and selected bivariate parametric families, we implement maximum likelihood methods to estimate parameters and calculate empirical dependence measures for each edge in each tree sequentially.

The function `XVineSeqEst` returns a list of three components:

1. A matrix of parameter estimates for each edge in each tree stored in `Params`.

2. A matrix of dependence measures stored in `DepMeasure`.

3. A matrix of effective sample sizes for each edge in each tree stored in `EffectSamp`.

In the code below, we use either multivariate inverted-Pareto samples from the limiting distribution or the rank-transformed data.

```{r, include=FALSE}
##  1. Use Pareto samples directly
#SeqEstOut=XVineSeqEst(data = Dat_P, Rank = FALSE, qt = 0.05, XVS=XVS, method = 'mle')
##  2. Use rank-based observations
SeqEstOut=XVineSeqEst(data = Dat_P, Rank = T, qt = 0.05, XVS=XVS, method = 'mle')
```

####  Compare the matrix of parameter estimates with the specified parameter values

```{r, echo=T}
round(SeqEstOut$Params,2) # Parameter estimates
round(XVS$pmat[,,1],2) # Parameter values
```
